---
title: "CALIBER code lists"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{caliber}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
bibliography: references.bib
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(ukbwranglrextra)
```

# Introduction

The CALIBER team have manually curated clinical code lists for 308 common health conditions, providing a rich resource for researchers working with electronic health records.[@kuan2019] All code lists are publicly available in csv format on github.[^1] These are divided into primary care (Read 2 codes and Medcodes) and secondary care (ICD10).

[^1]: See also the [HDRUK Phenotype Library](https://phenotypes.healthdatagateway.org/) for even more clinical code lists.

The UK Biobank contains both [linked primary and secondary care diagnostic records](https://biobank.ndph.ox.ac.uk/ukb/label.cgi?id=100091). Primary care records are in both Read 2 and Read 3 formats, while secondary care records are in both ICD10 and ICD9 formats (although the large majority of these are ICD10). Data analysts may therefore wish to extend the CALIBER resource by mapping to Read 3 and ICD9. The original Read 2 and ICD10 CALIBER codes also require reformatting to match the format in UK Biobank data.

The CALIBER repository may be imported into R, reformatted for use with UK Biobank data,[^2] and mapped to Read 3 and ICD9 equivalents using `map_caliber_for_ukb()`. This vignette outlines the steps performed to achieve this.

[^2]: See [UK Biobank resource 592](https://biobank.ndph.ox.ac.uk/ukb/refer.cgi?id=592) for further information on the various clinical coding systems.

# Reformatting Read 2 and ICD10 codes

## Read 2

-   Remove the last 2 digits from all Read 2 codes (these indicate whether the associated description is 'preferred' or 'alternative').
-   Filter for just one code description per code (in some cases, one or more alternative descriptions are included because they map to distinct Medcodes e.g. see preferred and alternative descriptions for '7K1D0' under ['Fracture of hip'](https://github.com/spiros/chronological-map-phenotypes/blob/master/primary_care/CPRD_fracture_hip.csv)).

The following read codes are not present in the Read 2 lookup table provided by UK Biobank resource 592:

| CALIBER disease                          | Unrecognised Read 2 code                                                                                                                          |
|------------------|------------------------------------------------------|
| Alcohol Problems                         | Z191.; Z1911; Z1912; Z4B1.                                                                                                                        |
| Anxiety disorders                        | Z481.; Z4L1.                                                                                                                                      |
| Coeliac disease                          | ZC2C2                                                                                                                                             |
| Crohn's disease                          | ZR3S.                                                                                                                                             |
| Dementia                                 | ZS7C5                                                                                                                                             |
| End stage renal disease                  | Z1A..; Z1A1.; Z1A2.; Z919.; Z9191; Z9192; Z9193; Z91A.                                                                                            |
| Erectile dysfunction                     | Z9E9.; ZG436                                                                                                                                      |
| Hearing loss                             | Z8B5.; Z8B51; Z8B53; Z8B55; Z911.; Z9111; Z9113; Z9114; Z9115; Z9117; Z9118; Z9119; Z911A; Z911B; Z911E; Z911G; Z9E81; ZE87.; ZL716; ZN569; ZN56A |
| Heart failure                            | ZRad.                                                                                                                                             |
| Intrauterine hypoxia                     | Z2648; Z2649; Z264A; Z264B                                                                                                                        |
| Lupus erythematosus (local and systemic) | ZRq8.; ZRq9.                                                                                                                                      |
| Obesity                                  | ZC2CM                                                                                                                                             |
| Other psychoactive substance misuse      | 9G24.; 9K4..; Z1Q62; Z416.                                                                                                                        |
| Tinnitus                                 | Z9112; ZEB..                                                                                                                                      |
| Transient ischaemic attack               | Z7CE7                                                                                                                                             |
| Urinary Incontinence                     | Z9EA.                                                                                                                                             |
| Visual impairment and blindness          | Z96..; Z961.; Z962.; ZK74.; ZN568; ZN56A; ZRhO.; ZRr6.                                                                                            |

## ICD10

-   Convert all ICD10 codes to the `ALT_CODE` format used in UK Biobank data. Note that while undivided three character ICD10 codes are flagged by an 'X' suffix in [UK Biobank resource 592](https://biobank.ndph.ox.ac.uk/ukb/refer.cgi?id=592 "Clinical coding classification systems and maps") (e.g. 'A38X', Scarlet fever), the suffix does not appear in the UK Biobank dataset itself (e.g. 'A38X' should instead appear as 'A38').
-   Expand all three character ICD10 codes to include their children. For example, 'D25' (Leiomyoma of uterus) is expanded to include 'D25', 'D250', 'D251', 'D252', 'D259'.[^3]
-   Expand any four character ICD10 codes with a 'modifier 5' to include their five character children. For example, 'M90.0' (Tuberculosis of bone) is expanded to include 'M900', 'M9000', 'M9001', 'M9002', 'M9003', 'M9004', 'M9005', 'M9006', 'M9007', 'M9008' and 'M9009'.

[^3]: See this [warning note](https://www.caliberresearch.org/portal/show/diabcomp_hes) (under the 'Coding lists' tab)

The following ICD10 codes are not present in the ICD10 lookup table provided by UK Biobank resource 592:[^4]

[^4]: These were retired after the 4th ICD10 edition, whereas the lookup table in UK Biobank resource 592 is based on the 5th edition.

| CALIBER disease                             | Unrecognised ICD10 code |
|---------------------------------------------|-------------------------|
| Infections of Other or unspecified organs   | A90; A91                |
| Viral diseases (excl chronic hepatitis/HIV) | A90; A91                |

## OPCS4

-   Remove '.' from all codes (e.g. 'H01.1' becomes 'H011').

# Mapping to Read 3 and ICD9

## Read 2 to Read 3

Mapping from Read 2 to Read 3 is performed using the `read_v2_read_ctv3` mapping sheet from [UK Biobank resource 592](https://biobank.ndph.ox.ac.uk/ukb/refer.cgi?id=592 "Clinical coding classification systems and maps"). Points to be aware of:

-   A minority of mappings in `read_v2_read_ctv3` are flagged as 'not assured' (`IS_ASSURED` '0'). These mappings are excluded by default.

## ICD10 to ICD9

Mapping from ICD10 to ICD9[^5] is performed using the `icd9_icd10` mapping sheet from [UK Biobank resource 592](https://biobank.ndph.ox.ac.uk/ukb/refer.cgi?id=592 "Clinical coding classification systems and maps"). Points to be aware of:

[^5]: Note that there are relatively few ICD9 diagnostic records.

-   There are a number of rows with missing values for either `DESCRIPTION_ICD9` or `DESCRIPTION_ICD10`, indicating that these codes have no ICD9/ICD10 equivalent.[^6]

-   One-to-many mappings occur in either direction (i.e. ICD9 to ICD10, and ICD10 to ICD9).

[^6]: Although some of these codes look like they should map to each other (e.g. ICD9 '0030' SALMONELLA GASTROENTERITIS and ICD10 'A020' Salmonella enteritis).

## Removing duplicates

# References
